Standardized crypto-loans on Cardano blockchain\thanks{Supported by Cardano Foundation.}

### 2021-02-03 papers/actus2021/article.tex
Actus article (#103)
citations

\begin{document}
\title{Standardized crypto-loans on Cardano blockchain\textquotedblright \thanks{Supported by organization x.}}

\maketitle
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here

\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and Second
Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
\authorrunning{F. Author et al.}

% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.

\institute{Princeton University, Princeton NJ 08544, USA \and Springer Heidelberg,
Tiergartenstr. 17, 69121 Heidelberg, Germany }
\email{lncs@springer.com}
\institute{\\
 \url{http://www.springer.com/gp/computer-science/lncs} \and ABC
Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
 }
\email{\{abc,lncs\}@uni-heidelberg.de}
\begin{abstract}
Crypto-loans are unique and innovative financial instruments that
allow trustless P2P lending, which could provide a safe and convenient
source of liquidity for cryptocurrency holders. In this article, we
explore a smart-contract framework for building standardized crypto-loans
with Marlowe DSL and Algorithmic Contract Types Unified Standards
at its core.

@@

A loan is a form of debt incurred by an individual or other entity.
The lender advances a sum of money to the borrower. In return, the
borrower agrees to a certain set of terms including any finance charges,
interest, repayment date, and other conditions.

@@

Counterparty might decide to reinvest profit received as interest
from the same loan. In the simplest case, this renders as compound
interest (TODO link). This can be modeled through interest accrual
and capitalisation, for instance contracts from the ACTUS specification
accrue interest between interest payments and can transfer interest
to a notional during interest capitalisation event (IPCL).

@@

Marlowe is a high-level, domain-specific language (DSL) for writing
financial contracts on blockchain. Marlowe is defined by an executable
semantics in Haskell, and has been implemented on the UTxO-based Cardano
blockchain.

The main advantage of using Marlowe to carry ACTUS logic is enhanced
security that static analysis {[}TODO link{]} provides. SMT-solvers
gain increasing popularity as an alternative to fuzzing when it comes
to testing software for vulnerabilities {[}TODO link{]}. Smart contracts
generated from ACTUS specification are quite complicated, so manually
testing every possible execution path for every possible combination
of contract terms is not feasible. Static analysis, in turn, can reduce
the amount of ``dumb'' mistakes that could cost millions to the
users by ensuring obvious properties of a contract. It is not a panacea,
however, unlike with fuzzing, if SA says that certain property holds
- it does it with 100\% assurance.

@@

ACTUS defines the logic embedded in legal agreements that eventually
turn the contract terms into actual cash flows, or more generally
business events. Most of Its basic contract types represent different
variations of lending contracts.

@@

This is due to the risk factor's state not being known at instantiation
time. In the case of Cardano blockchain, these values are usually
provided through an ``Oracle'' mechanism. Oracle could be a trusted
party providing necessary data or network of parties under consensus
{[}find a quote{]}. From Marlowe DSL perspective, the exact mechanism
that provides external data is less important, as Marlowe abstracts
over IO by requiring particular type of input (``Choice'') protected
with a cryptographic signature (Choice owner). As a result, the event
of receiving data from Oracle is treated the same as receiving integer
numbers from user input in other languages

@@

The ACTUS standard is specified in terms of {[}link{]} scheduling,
payoff and state transition functions that are polymorphic on event
and contract type. The specification also follows quite specific naming
conventions that are incompatible with Haskell's conventions. The
executable specification follows original ACTUS conventions as close
as possible in order to ease code base maintenance when faced with
updates of original ACTUS spec (link to github).

@@

\begin{figure}
\includegraphics[width=1\textwidth]{modules} \caption{Modules responsible for contract generation.}
\label{fig1-1} 
\end{figure}

@@

where $receiveData$ - is asking an Oracle for Marlowe Choice if needed,
$calculatePayoff$ - calculates payoff formula. For fixed-rate contracts
this is optimised into a precalculated constant. $processPayoff$
- awaits Deposit of payoff amount from a party: if deposit is transferred
it redirects the funds to a counterparty - otherwise it transfers
the collateral to a counterparty and closes the contract. The chain
is generated from the fixed (known in advance) schedule of events
using $SCHED()$function from ACTUS.

\begin{figure}
\includegraphics[width=1\textwidth]{flowchart} \caption{Chain of sub-contracts representing ACTUS logic}
\label{fig1-1-1} 
\end{figure}

@@

Marlowe doesn't allow perpetual contracts even if their recursion
is productive (example: perpetual swap) - thus we can't support certain
contract types from ACTUS specification, namely the ones that don't
have a defined maturity date (like UMP). In a long-term perspective
- there is a possible workaround: contracts with no maturity date
could be represented as actors with finite amounts of state transitions.
We prototyped this approach in a separate branch, however it does
seem more prone to errors comparing to rendering predefined schedules
and more importantly it greatly affects static analysis as the amount
of ``reduction steps'' grows from $N_{scheduled}=count(event)$
to $N_{stateTransitions}=\frac{max(date(event))-min(date(event))}{precision}$.
See \href{https://github.com/dk14/plutus/tree/ACTUS}{https://github.com/dk14/plutus/tree/ACTUS}

@@

\begin{figure} 
\includegraphics[width=0.8\textwidth,natwidth=664,natheight=552]{labs.png}
\caption{Actus Labs - online tool for generating Actus contracts for Marlowe.} 
\label{fig2} 
\end{figure}

@@

QuickCheck contract terms generators also allow to ensure arbitrary
properties of a Marlowe contract. This could be enhanced with Marlowe's
static analysis feature by utilising Assert operator:

@@

This implicit refund is easy to overlook by smart-contract developers
as plain Close is often used as a default action in case of unexpected
behaviour like timeouts(especially Choice timeout). This can easily
lead to multi-million dollar mistakes if malicious Alice decides to
exploit an auto-refund feature in order to get her collateral without
paying back. Here's the example of simple attack vector: 
\begin{enumerate}
\item Alice creates a contract where Deposit timeout would lead to Close 
\item Bob doesn't know or test the timeout path. Even if Bob is a programmer
- he might not be aware that Closing contract would cause collateral
refund 
\item Alice puts collateral and gets notional from Bob 
\item Alice doesn't pay for a loan (Deposit timeout) 
\item Alice gets her collateral back 
\item Bob loses notional 
\end{enumerate}
In order to prevent this from happening, an additional ''Auto-Refund''
security check was introduced as part of Static Analysis {[}TODO link{]}
tooling which notifies users about all ``Close'' constructs that
can lead to automatic refunds and encourages users to write explicit
logic for edge cases.

@@

Unlike current popular DeFi lending approaches (like Harvest Finance)
- Marlowe ACTUS-based implementation tends to rely on ``good old''
matching instead of pooling. While pooling is proven to be superior
to order-book based approaches when it comes to automated market makers,
using it for lending has been shown to be more susceptible to attacks
(for example ``\href{https://www.coindesk.com/harvest-finance-24m-attack-triggers-570m-bank-run-in-latest-defi-exploit}{flash loan}'').

>>>>

\begin{document}
\title{Standardized crypto-loans on Cardano blockchain\thanks{Supported by Cardano Foundation.}}
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Crypto-loans are unique and innovative financial instruments that
allow trustless P2P lending, which could provide a safe and convenient
source of liquidity for cryptocurrency holders. In this article, we
explore a smart-contract framework for building standardized crypto-loans
with Marlowe DSL and Algorithmic Contract Types Unified Standards
at its core.

@@

A loan is a form of debt incurred by an individual or other entity.
The lender advances a sum of money to the borrower. In return, the
borrower agrees to a certain set of terms including any finance charges,
interest, repayment date, and other conditions \cite{loan}.

@@

Counterparty might decide to reinvest profit received as interest
from the same loan. In the simplest case, this renders as compound
interest. This can be modeled through interest accrual and capitalisation,
for instance contracts from the ACTUS specification accrue interest
between interest payments and can transfer interest to a notional
during interest capitalisation event (IPCL).

@@

Marlowe is a high-level, domain-specific language (DSL) for writing
financial contracts on blockchain\cite{marlowe}. Marlowe is defined
by an executable semantics in Haskell, and has been implemented on
the UTxO-based Cardano blockchain.

The main advantage of using Marlowe to carry ACTUS logic is enhanced
security that static analysis provides. SMT-solvers gain increasing
popularity as an alternative to fuzzing when it comes to testing software
for vulnerabilities\cite{smt}. Smart contracts generated from ACTUS
specification are quite complicated, so manually testing every possible
execution path for every possible combination of contract terms is
not feasible. Static analysis, in turn, can reduce the amount of ``dumb''
mistakes that could cost millions to the users by ensuring obvious
properties of a contract. It is not a panacea, however, unlike with
fuzzing, if SA says that certain property holds - it does it with
100\% assurance.

@@

ACTUS defines the logic embedded in legal agreements that eventually
turn the contract terms into actual cash flows, or more generally
business events \cite{actus}. Most of its basic contract types represent
different variations of lending contracts.

@@

This is due to the risk factor's state not being known at instantiation
time. In the case of Cardano blockchain, these values are usually
provided through an ``Oracle'' mechanism\cite{oracles}. Oracle
could be a trusted party providing necessary data or network of parties
under consensus {[}find a quote{]}. From Marlowe DSL perspective,
the exact mechanism that provides external data is less important,
as Marlowe abstracts over IO by requiring particular type of input
(``Choice'') protected with a cryptographic signature (Choice owner).
As a result, the event of receiving data from Oracle is treated the
same as receiving integer numbers from user input in other languages

@@

The ACTUS standard is specified in terms of scheduling, payoff and
state transition functions that are polymorphic on event and contract
type. The specification also follows quite specific naming conventions
that are incompatible with Haskell's conventions. The executable specification
follows original ACTUS conventions as close as possible in order to
ease code base maintenance when faced with updates of original ACTUS
spec (link to github).

@@

\begin{figure}
\includegraphics[width=1\textwidth]{modules} \caption{Modules responsible for contract generation.}
\label{fig1} 
\end{figure}

@@

where $receiveData$ - is asking an oracle for Marlowe Choice if needed,
$calculatePayoff$ - calculates payoff formula. For fixed-rate contracts
this is optimised into a precalculated constant. $processPayoff$
- awaits Deposit of payoff amount from a party: if deposit is transferred
it redirects the funds to a counterparty - otherwise it transfers
the collateral to a counterparty and closes the contract. The chain
is generated from the fixed (known in advance) schedule of events
using $SCHED()$function from ACTUS.

\begin{figure}
\includegraphics[width=1\textwidth]{flowchart} \caption{Chain of sub-contracts representing ACTUS logic}
\label{fig2} 
\end{figure}

@@

Marlowe doesn't allow perpetual contracts even if their recursion
is productive (example: perpetual swap) - thus we can't support certain
contract types from ACTUS specification, namely the ones that don't
have a defined maturity date (like UMP). In a long-term perspective
- there is a possible workaround: contracts with no maturity date
could be represented as actors with finite amounts of state transitions.
We prototyped this approach in a separate branch, however it does
seem more prone to errors comparing to rendering predefined schedules
and more importantly it greatly affects static analysis as the amount
of ``reduction steps'' grows from $N_{scheduled}=count(event)$
to $N_{stateTransitions}=\frac{max(date(event))-min(date(event))}{precision}$.

@@

\begin{figure} 
\includegraphics[width=0.8\textwidth,natwidth=664,natheight=552]{labs.png}
\caption{Actus Labs - online tool for generating Actus contracts for Marlowe.} 
\label{fig3} 
\end{figure}

@@

QuickCheck\cite{qc} contract terms generators also allow to ensure
arbitrary properties of a Marlowe contract. This could be enhanced
with Marlowe's static analysis feature by utilising Assert operator:

@@

This implicit refund is easy to overlook by smart-contract developers
as plain Close is often used as a default action in case of unexpected
behaviour like timeouts(especially Choice timeout). This can easily
lead to multi-million dollar mistakes if malicious Alice decides to
exploit an auto-refund feature in order to get her collateral without
paying back. Here's the example of simple attack vector: 
\begin{enumerate}
\item Alice creates a contract where Deposit timeout would lead to Close 
\item Bob doesn't know or test the timeout path. Even if Bob is a programmer
- he might not be aware that Closing contract would cause collateral
refund 
\item Alice puts collateral and gets notional from Bob 
\item Alice doesn't pay for a loan (Deposit timeout) 
\item Alice gets her collateral back 
\item Bob loses notional 
\end{enumerate}
In order to prevent this from happening, an additional ''Auto-Refund''
security check was introduced as part of Static Analysis tooling which
notifies users about all ``Close'' constructs that can lead to automatic
refunds and encourages users to write explicit logic for edge cases.

@@

Unlike current popular DeFi lending approaches - Marlowe ACTUS-based
implementation tends to rely on trade-matching instead of pooling.
While asset pooling is proven to be superior to order-book based approaches
when it comes to automated market makers, using it for lending has
been shown to be more susceptible to attacks\cite{flash-loan}.
